### 2014-10-30 fundamentals/intro.tex
updated minor edits

The \Cyclus Nuclear Fuel Cycle Simulator framework and 
its modeling ecosystem answer this need for a next generation nuclear fuel cycle 
simulator. Dynamic simulation of nuclear fuel cycles with discrete agents 
and up to isotopic resolution support 
levels of detail appropriate for a varied community of analysts
interested in a range of analysis types. Additionally \Cyclus employs an 
open development process that supports user access and 
encourages developer extensions.

Historically, nuclear fuel cycle simulators have estimated 
nuclear fuel cycle performance using software 
ranging from ad hoc spreadsheet-driven flow calculators to highly specialized 
proprietary models. To date, current tools are either static, fleet-based, 
privately distributed, or all of the above. Each of these qualities present a 
severe challenge to modeling fidelity, generality, efficiency, robustness, and 
scientific transparency.

An alternative to these approaches is a dynamic, agent-based simulator with an 
open platform and an ecosystem of research-driven capabilities.  Modern 
insight indicates that dynamic nuclear fuel cycle analysis more realistically 
supports a broader range of simulation goals than static analysis 
\cite{piet_dynamic_2011}. Similarly, experience in the greater field of systems 
analysis indicates that agent-based modeling enables more flexible simulation 
control, without loss of generality \cite{macal_agent-based_2010}. Finally, openness 
allows cross-institutional collaboration, increases code robustness, and 
enables the cultivation of an ecosystem of calculation options, 
\cite{softwarecarpentryresource}.  The fundamental concepts of the \Cyclus 
nuclear fuel cycle simulator capture these modern insights so that challenges 
in nuclear fuel cycle analysis can be better addressed.

@@

Historically, the national laboratories have driven development and regulated 
the use of their own tools such as \gls{VISION}\cite{jacobson_verifiable_2010}, 
\gls{DANESS}\cite{van_den_durpel_daness_2009}, 
\gls{DYMOND}\cite{yacout_modeling_2005}, and 
\gls{NFCSim}\cite{schneider_nfcsim:_2005}. 
Outside of the national laboratories, researchers create new nuclear fuel cycle 
simulation tools when existing tools are not available or not sufficiently 
general to calculate their metrics of interest with their method of choice.  
With limited access to the national laboratory tools and a need to customize 
them for research purposes, universities and private industry researchers 
``reinvent the wheel'' by developing tools of their own from scratch and 
tailored to their own needs (e.g., \gls{CAFCA}\cite{guerin_impact_2009} and 
\gls{COSI}\cite{boucher_cosi_2005,boucher_cosi:_2006,meyer_new_2009,coquelet-pascal_comparison_2011} respectively).

This cycle has resulted in a proliferation of specialized
fuel cycle simulators for the calculation of fuel cycle metrics.  Accordingly, methods 
for calculating those metrics vary among simulators. Some model the 
system of facilities, economics, and materials at static equilibrium, while 
others capture its dynamics.  
Similarly, while some simulators discretely model material and facilities, 
others treat materials in streams and facilities in fleets, in an aggregated 
fashion. Depending on their intended use, simulators often overspecialize, 
modeling a single aspect of the fuel cycle in great detail while neglecting 
others. A simulator created for policy modeling might have excellent 
capability in economics while material isotopics and transmutation detail are 
neglected.

The \Cyclus simulator emerged from a line of tools seeking to break that cycle.  
An \gls{LDRD} project begun at \gls{INL}, 
\gls{GENIUSv1}\cite{dunzik-gougar_global_2007,jain_transitioning_2006} sought 
to provide generic regional capability.  Based on lessons learned from 
\gls{GENIUSv1}, the 
\gls{GENIUSv2}\cite{oliver_studying_2009,huff_geniusv2_2009} code sought to 
provide more generality and an extensible interface to facilitate 
collaboration.  The \Cyclus project, in turn, improved upon the \gls{GENIUSv2} 
project by implementing increased modularity and encapsulation.  The result is  
a dynamic simulator that treats both materials and facilities discretely, with 
an architecture that permits simulations with multiple and variable levels of 
fidelity. Its agent-based framework for simulation and tracking of 
transformation and trade of resources between facilities within institutions 
and regions greatly improves upon the existing field of fuel cycle simulators.

@@

Furthermore, modeling the evolution of a physics-dependent, international nuclear fuel supply
chain is a multiscale problem, but existing tools are not sufficiently 
multi-scale. Tools have either focused on macro effects, e.g., the fleet-level
stocks and flows of commodities, or micro effects, e.g., the used fuel
composition fast reactor fuel. Accordingly, each focus has driven the
development of these tools, rendering the task of answering questions between
the macro and micro levels challenging within a single tool. 
However, an extensible architecture and discrete object tracking allow 
the creation and exchange ability of custom \Cyclus archetypes at any level of fidelity.

@@

First, analysts across research institutions can take advantage of the 
simulation logic \gls{API} and archetype ecosystem when they apply \Cyclus to 
their specific problem. This extensibility, magnified by an open source 
development process, frees the community of users to contribute desired 
features independen from the team of \Cyclus core developers. This 
mend-and-extend paradigm, common in modern software development, reduces 
maintenance overhead and improves code robustness \cite{citation_needed}. 
Finally, the researcher can therefore focus on creating or customizing nuclear 
facility, institution, resource, and toolkit models within their specific area 
of technical expertise.

@@

We expect that there are a number of types of users and developers who will 
interact with the \Cyclus framework. National laboratory researchers,
academic researchers, private-industry economic forcasters, and even students. 
in engineering courses will benefit from the ease of development and ease of 
customization provided by the modular architecture of \Cyclus.

@@

\Cyclus utilizes an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and it takes advantage of modern computing resources.  
Furthermore, agent-based modeling provides an analyst a more natural 
perspective from which to model a given problem. In the nuclear fuel cycle 
context for example, an analyst can design an agent with reactor-related logic 
entirely independently from a separate agent with fuel fabrication-related 
logic. Defining the behavior of both agents according to the 
\gls{API} contract is sufficient for them to interact with one another as 
bonafide actors in the simulation.  The two archetype libraries, therefore, need 
not have any shared knowledge of one another for them to be building blocks in 
the same simulation.

System dynamics is a popular approach for modeling nuclear fuel cycles in current
tools 
\cite{jacobson_vision_2009,van_den_durpel_daness_2009,guerin_impact_2009,guerin_benchmark_2009}. However, it has been shown in systems analysis literature
\cite{macal_agent-based_2010} that, formally, the
set of system dynamics models is a strict subset of agent-based models.
That is, any system dynamics model can be translated
into an agent-based model. Furthermore, agent-based
techniques can provide a richer model than can system dynamics. To enable the 
broadest range of simulations in a generic fashion, \gls{ABM} is therefore 
superior to system dynamics.

>>>>

The \Cyclus Nuclear Fuel Cycle Simulator framework and 
its modeling ecosystem answer this need for a next generation nuclear fuel cycle 
simulator. Dynamic simulation of nuclear fuel cycles with discrete agents 
and per nuclide species resolution support 
levels of detail appropriate for a varied community of analysts
interested in a range of analysis types. Additionally \Cyclus employs an 
open development process that supports user access and 
encourages developer extensions.

Historically, nuclear fuel cycle simulators have estimated 
nuclear fuel cycle performance using software 
ranging from ad hoc spreadsheet-driven flow calculators to highly specialized 
system dynamics
proprietary models. To date, current tools are either static, fleet-based, 
privately distributed, or all of the above. Each of these qualities present a 
severe challenge to modeling fidelity, generality, efficiency, robustness, and 
scientific transparency.

An alternative to these approaches is a dynamic, agent-based simulator with an 
open platform and an ecosystem of research-driven capabilities.  Modern 
insight indicates that dynamic nuclear fuel cycle analysis more realistically 
supports a broader range of simulation goals than static analysis 
\cite{piet_dynamic_2011}. Similarly, experience in the broader field of systems 
analysis indicates that agent-based modeling enables more flexible simulation 
control, without loss of generality \cite{macal_agent-based_2010}. Finally, openness 
allows cross-institutional collaboration, increases code robustness, and 
enables the cultivation of an ecosystem of calculation options, 
\cite{softwarecarpentryresource}.  The fundamental concepts of the \Cyclus 
nuclear fuel cycle simulator capture these modern insights so that challenges 
in nuclear fuel cycle analysis can be better addressed.

@@

Historically, the national laboratories have driven development and regulated 
the use of their own tools such as \gls{VISION}\cite{jacobson_verifiable_2010}, 
\gls{DANESS}\cite{van_den_durpel_daness_2009}, 
\gls{DYMOND}\cite{yacout_modeling_2005}, and 
\gls{NFCSim}\cite{schneider_nfcsim:_2005}. 
Outside of the national laboratories, researchers create new nuclear fuel cycle 
simulation tools when existing tools are not available or not sufficiently 
general to calculate their metrics of interest with their method of choice.  
With limited access to the national laboratory tools and a need to customize 
them for research purposes, universities and private industry researchers 
``reinvent the wheel'' by developing tools of their own from scratch and 
tailored to their own needs, for example \gls{CAFCA}\cite{guerin_impact_2009} and 
\gls{COSI}\cite{boucher_cosi_2005,boucher_cosi:_2006,meyer_new_2009,coquelet-pascal_comparison_2011} respectively.

This cycle has resulted in a proliferation of specialized
fuel cycle simulators for the calculation of fuel cycle metrics.  Accordingly, methods 
for calculating those metrics vary among simulators. Some model the 
system of facilities, economics, and materials at static equilibrium. Meanwhile 
other simulators capture its dynamics.  
Similarly, while some simulators discretely model material and facilities, 
others treat materials in streams and facilities in fleets, in an aggregated 
fashion. Depending on their intended use, simulators often overspecialize, 
modeling a single aspect of the fuel cycle in great detail while neglecting 
others. A simulator created for policy modeling might have excellent 
capability in economics while material isotopics and transmutation detail are 
neglected.

The \Cyclus simulator emerged from a line of tools seeking to break that cycle.  
An \gls{LDRD} project begun at \gls{INL}, 
\gls{GENIUSv1}\cite{dunzik-gougar_global_2007,jain_transitioning_2006} sought 
to provide generic regional capability.  Based on lessons learned from 
\gls{GENIUSv1}, the 
\gls{GENIUSv2}\cite{oliver_studying_2009,huff_geniusv2_2009} code sought to 
provide more generality and an extensible interface to facilitate 
collaboration.  The \Cyclus project, in turn, improved upon the \gls{GENIUSv2} 
effort by implementing increased modularity and encapsulation.  The result is  
a dynamic simulator that treats both materials and facilities discretely, with 
an architecture that permits simulations with multiple and variable levels of 
fidelity. Its agent-based framework for simulation and tracking of 
transformation and trade of resources between facilities within institutions 
and regions greatly improves upon the existing field of fuel cycle simulators.

@@

Furthermore, modeling the evolution of a physics-dependent, international nuclear fuel supply
chain is a multiscale problem, but existing tools are not sufficiently 
multiscale. Tools have either focused on macro effects, e.g., the fleet-level
stocks and flows of commodities, or micro effects, e.g., the used fuel
composition fast reactor fuel. Accordingly, each focus has driven the
development of these tools, rendering the task of answering questions between
the macro and micro levels challenging within a single tool. 
However, an extensible architecture and discrete object tracking allow 
the creation and exchange ability of custom \Cyclus archetypes at any level of fidelity.

@@

First, analysts across research institutions can take advantage of the 
simulation logic \gls{API} and archetype ecosystem when they apply \Cyclus to 
their specific problem. This extensibility, magnified by an open source 
development process, frees the community of users to contribute desired 
features independen from the team of \Cyclus core developers. This 
mend-and-extend paradigm, common in modern software development, reduces 
maintenance overhead and improves code robustness \cite{citation_needed}. 
The researcher can therefore focus on creating or customizing nuclear 
facility, institution, resource, and toolkit models within their specific area 
of technical expertise.

@@

We expect that there are a number of types of users and developers who will 
interact with the \Cyclus framework: national laboratory researchers,
academic researchers, private-industry economic forcasters, and students
in engineering courses will benefit from the ease of development and ease of 
customization provided by the modular architecture of \Cyclus.

@@

\Cyclus implements an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and it takes advantage of modern computing resources.  
Furthermore, agent-based modeling provides an analyst a more natural 
perspective from which to model a given problem. In the nuclear fuel cycle 
context for example, an analyst can design an agent with reactor-related logic 
entirely independently from a separate agent with fuel fabrication-related 
logic. Defining the behavior of both agents according to the 
\gls{API} contract is sufficient for them to interact with one another as 
bonafide actors in the simulation.  The two archetypes, therefore, need 
not have any shared knowledge of one another for them to be building blocks in 
the same simulation.

System dynamics is a popular approach for modeling nuclear fuel cycles 
\cite{jacobson_vision_2009,van_den_durpel_daness_2009,guerin_impact_2009,guerin_benchmark_2009}. However, it has been shown in systems analysis literature
\cite{macal_agent-based_2010} that, formally, the
set of system dynamics models is a strict subset of agent-based models.
That is, any system dynamics model can be translated
into an agent-based model. Furthermore, agent-based
techniques can provide a richer model than can system dynamics. To enable the 
broadest range of simulations in a generic fashion, \gls{ABM} is therefore 
superior to system dynamics.

### 2014-10-30 fundamentals/qa.tex
updated minor edits

It is important to note that before a proposed
code changes is allowed into main stream cyclus, all of the tests must pass
and the change must be covered by a test, either new or existing. This is 
matter of policy for the developers. Additionally, test code is equally likely 
as the code it is testing to contain errors. To prevent an infinite recurrsion 
of testing-the-tests, it is assumed that dilligence and one level of testing 
is sufficient to meet the aims of the quality assurance.

@@

Intergration tests are any tests that combine multiple elemnents of the 
cyclus interface and test that they work correctly with each other.  By analogy, 
simply because the gears (units) are made correctly does not imply that the 
clock (integration) will run smoothly, run at all, or give the correct time.
In cyclus and cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of 
time. Precanned input files have thus been contsructed. Their results 
are inspected and compared via nose \cite{pellerin_nose_2007}, a Python test framework.
Python will run cyclus as a subprocess for each integration test. In this
way cyclus code units are tested in the full context that they will
executed. This category of testing is especially useful for ensuring that 
major cyclus components are functioning as expected.

@@

Regression tests are tests which ensure that significant changes do not 
occur over the course of cyclus development. Such a change is called a 
\emph{regression} because the new version is almost always wrong.
Regression tests are implemeneted similarly to integration tests.
Nose is used to execute fully fledged cyclus simulations whose results
are then compared. In this category however the comparison is done against 
the output of the same input file when run with a previous version of cyclus, 
typically the last released version.
In some sense, regression tests are `dumb' in that they do 
not care about the contents of a simulation being correct, only whether or not 
it changed. They thus pick up glaring time-sensitive mistakes that may 
not appear elsewhere though they do not purport to make claims about the
quality of the physics modeled. The cycamore project is the primary location of
regression tests. This is because the archetypes here are sophisticated enough
to make interesting input files. Only a handful ($<10$) time steps are needed 
for regression testing.

@@

All of the public interface (the API) must be documented as part of cyclus
QA policy. This certifies that the intention for a code unit is communicated 
in way that is separate from its implementation. This policy is important 
because implementations are often wrong, hence bugs. They are thus unreliable 
resources about themselves. Truly trustworthy communication about what 
software should be doing comes from the author explicitly stating the 
goals of an API in prose form. Furure development of the code is then able to 
compare the implemetation versus the documentation to discern whether or not they 
are in agreement. Documentation is thus a secondary higher-level information 
stream.  In cyclus, this information is aggregated together into static 
websites with the doxygen \cite{van_heesch_doxygen:_2008} and sphinx 
\cite{brandl_sphinx_2014} 
tools. These websites are placed online at \url{http://fuelcycle.org}.

@@

Version control is the history-preserving mechanism by which the provenance of 
a code is recored. Cyclus uses a well-established tool called git 
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utilty in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every 
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there and many praises and criticisms of git, one 
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simeltaneous
branches may exist at all point in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an 
accompanying message. Thus 
it is possible for cyclus to accurarely replay the entirety of who did what to the
code when.

Cyclus uses a strategy known as \emph{git flow} 
\cite{kalliamvakou_code-centric_2014} 
to manage topical branches where individuals develop code, the mainline development
branch (simply known as develop), and the stable branch (called master).
In addition to these rules and delineation, cyclus also imposes rules on 
when it is accecptable from a quality control standpoint to merge code from 
the personal topical branches into develop. All software at this stage 
must have sufficient and passing tests and reasonable documentation. Without
these core pieces proposed changes are not allowed into cyclus.

@@

Continuous integration (CI) the idea that software should be tested and validated 
as it is being developed, rather than as a final stage in a longer development 
cycle.  Under CI every pull request is tested independently immeadiately after 
it is proposed. Such tests shoudl be run on all officially supported platforms. 
Cyclus uses a CI platform called Polyphemus \cite{scopatz_polyphemus_2014}.

>>>>

It is important to note that before a proposed
code change is allowed into mainline cyclus, all of the tests must pass
and the change must be covered by a test, either new or existing. This is a
matter of policy for the developers. Additionally, test code is equally likely 
as the code it is testing to contain errors. To prevent an infinite recurrsion 
of testing-the-tests, it is assumed that dilligence and one level of testing 
is sufficient to meet the aims of the quality assurance.

@@

Intergration tests are any tests that combine multiple elements of the 
cyclus interface and test that they work correctly with each other.  By analogy, 
simply because the gears (units) are made correctly does not imply that the 
clock (integration) will run smoothly, run at all, or give the correct time.
In cyclus and cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of 
time. Precanned input files have thus been contsructed. Their results 
are inspected and compared via nose \cite{pellerin_nose_2007}, a Python test framework.
Python will run cyclus as a subprocess for each integration test. In this
way cyclus code units are tested in the full context that they will be
executed. This category of testing is especially useful for ensuring that 
major cyclus components are functioning as expected.

@@

Regression tests are tests which ensure that significant changes do not 
occur over the course of cyclus development. Such a change is called a 
\emph{regression} because the new version is almost always wrong version in the 
event of a difference.
Regression tests are implemeneted similarly to integration tests.
Nose is used to execute fully fledged cyclus simulations whose results
are then compared. In this category however the comparison is done against 
the output of the same input file when run with a previous version of cyclus, 
typically the last released version.
In some sense, regression tests are `dumb' in that they do 
not care about the contents of a simulation being correct, only whether or not 
it changed. They thus pick up glaring time-sensitive mistakes that may 
not appear elsewhere though they do not purport to make claims about the
quality of the physics modeled. The cycamore project is the primary location of
regression tests. This is because the archetypes here are sophisticated enough
to make interesting input files. Only a handful ($<10$) time steps are needed 
for regression testing.

@@

All of the public interfaces must be documented as part of cyclus
QA policy. This certifies that the intention for a code unit is communicated 
in way that is separate from its implementation. This policy is important 
because implementations are often wrong. Software implementations are unreliable 
resources about themselves. Truly trustworthy communication about what 
software should be doing comes from the author explicitly stating the 
goals of an API in prose form. Furure development of the code is then able to 
compare the implemetation versus the documentation to discern whether or not they 
are in agreement. Documentation is thus a secondary higher-level information 
source.  In cyclus, this information is aggregated together into static 
websites with the doxygen \cite{van_heesch_doxygen:_2008} and sphinx 
\cite{brandl_sphinx_2014} 
tools. These websites are placed online at \url{http://fuelcycle.org}.

@@

Version control is the history-preserving mechanism by which the provenance of 
a code is recored. Cyclus uses a well-established tool called git 
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utilty in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every 
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there and many praises and criticisms of git, one 
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simeltaneous
branches may exist at all points in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an 
accompanying message. Thus 
it is possible for cyclus to accurarely replay the entirety of who did what to the
code when.

Cyclus uses a strategy known as \emph{git flow} 
\cite{kalliamvakou_code-centric_2014} 
to manage topical branches where individuals develop code, the mainline development
branch (simply known as develop), and the stable branch (called master).
In addition to these rules and delineation, cyclus also imposes rules on 
when it is accecptable from a quality control standpoint to merge code from 
the personal topical branches into develop. All software at this stage 
must have sufficient and passing tests and reasonable documentation. Without
these core pieces, proposed changes are not allowed into cyclus.

@@

Continuous integration (CI) is the idea that software should be tested and validated 
as it is being developed, rather than as a final stage in a longer development 
cycle.  Under CI every pull request is tested independently immeadiately after 
it is proposed. Such tests shoudl be run on all officially supported platforms. 
Cyclus uses a CI platform called Polyphemus \cite{scopatz_polyphemus_2014}.
